\bibliographystyle{abbrv}
\begin{thebibliography}{10}

\bibitem{bcm4751}
{Broadcom BCM4751 Integrated Monolithic GPS Receiver}.
\newblock
  \url{http://www.broadcom.com/products/GPS/GPS-Silicon-Solutions/BCM4751}.

\bibitem{livelabs-url}
Livelabs.
\newblock \url{http://centres.smu.edu.sg/livelabs/}.

\bibitem{avetisyan:computer:2010}
A.~I. Avetisyan, R.~Campbell, I.~Gupta, M.~T. Heath, S.~Y. Ko, G.~R. Ganger,
  M.~A. Kozuch, D.~O'Hallaron, M.~Kunze, T.~T. Kwan, K.~Lai, M.~Lyons, D.~S.
  Milojicic, H.~Y. Lee, Y.~C. Soh, N.~K. Ming, J.-Y. Luke, and H.~Namgoong.
\newblock {Open Cirrus: A Global Cloud Computing Testbed}.
\newblock {\em IEEE Computer}, 43(4):35--43, Apr. 2010.

\bibitem{banerjee:ubicomp:2007}
N.~Banerjee, A.~Rahmati, M.~D. Corner, S.~Rollins, and L.~Zhong.
\newblock {Users and Batteries: Interactions and Adaptive Energy Management in
  Mobile Systems}.
\newblock In {\em Proceedings of the 9th International Conference on Ubiquitous
  Computing (UbiComp)}, 2007.

\bibitem{google-total-activations}
{BGR Media, LLC}.
\newblock {Google announces 1 billion Android activations, teases Android 4.4}.
\newblock \url{http://goo.gl/u8m4HA}, September 2013.

\bibitem{hao:sensys:2013}
T.~Hao, G.~Xing, and G.~Zhou.
\newblock {iSleep: Unobtrusive Sleep Quality Monitoring using Smartphones}.
\newblock In {\em Proceedings of the 8th ACM Conference on Embedded Networked
  Sensor Systems (SenSys)}, 2013.

\bibitem{idc-smartphone-growth}
{International Data Corporation}.
\newblock {Worldwide Mobile Phone Growth Expected to Drop to 1.4% in 2012
  Despite Continued Growth Of Smartphones}.
\newblock \url{http://goo.gl/ROLY0}.

\bibitem{humanmobility-mobisys12}
S.~Isaacman, R.~Becker, R.~C\'{a}ceres, M.~Martonosi, J.~Rowland,
  A.~Varshavsky, and W.~Willinger.
\newblock Human mobility modeling at metropolitan scales.
\newblock In {\em Proceedings of the 10th international conference on Mobile
  systems, applications, and services}, MobiSys '12, pages 239--252, New York,
  NY, USA, 2012. ACM.

\bibitem{ace-mobisys12}
S.~Nath.
\newblock Ace: exploiting correlation for energy-efficient and continuous
  context sensing.
\newblock In {\em Proceedings of the 10th international conference on Mobile
  systems, applications, and services}, MobiSys '12, pages 29--42, New York,
  NY, USA, 2012. ACM.

\bibitem{peterson:ccr:2003}
L.~Peterson, T.~Anderson, D.~Culler, and T.~Roscoe.
\newblock {A Blueprint for Introducing Disruptive Technology into the
  Internet}.
\newblock {\em SIGCOMM Comput. Commun. Rev.}, 33(1):59--64, Jan. 2003.

\bibitem{raychaudhuri:tridentcom:2005}
D.~Raychaudhuri, M.~Ott, and I.~Secker.
\newblock {ORBIT Radio Grid Tested for Evaluation of Next-Generation Wireless
  Network Protocols}.
\newblock In {\em Proceedings of the First International Conference on Testbeds
  and Research Infrastructures for the DEvelopment of NeTworks andA COMmunities
  (TRIDENTCOM)}, 2005.

\bibitem{shye:micro:2009}
A.~Shye, B.~Scholbrock, and G.~Memik.
\newblock {Into the Wild: Studying Real User Activity Patterns to Guide Power
  Optimizations for Mobile Architectures}.
\newblock In {\em Proceedings of the 42nd Annual IEEE/ACM International
  Symposium on Microarchitecture (MICRO)}, 2009.

\bibitem{netsense-hotplanet}
A.~Striegel, S.~Liu, L.~Meng, C.~Poellabauer, D.~Hachen, and O.~Lizardo.
\newblock Lessons learned from the netsense smartphone study.
\newblock In {\em Proceedings of the 5th ACM workshop on HotPlanet}, HotPlanet
  '13, pages 51--56, New York, NY, USA, 2013. ACM.

\bibitem{nowar-mobisys12}
H.~Wang, S.~Sen, A.~Elgohary, M.~Farid, M.~Youssef, and R.~R. Choudhury.
\newblock No need to war-drive: unsupervised indoor localization.
\newblock In {\em Proceedings of the 10th international conference on Mobile
  systems, applications, and services}, MobiSys '12, pages 197--210, New York,
  NY, USA, 2012. ACM.

\bibitem{wang:mobisys:2013}
Y.~Wang, J.~Yang, H.~Liu, Y.~Chen, M.~Gruteser, and R.~P. Martin.
\newblock {Sensing Vehicle Dynamics for Determining Driver Phone Use}.
\newblock In {\em Proceeding of the 11th annual international conference on
  Mobile systems, applications, and services}, MobiSys '13, 2013.

\bibitem{werner-allen:ipsn:2005}
G.~Werner-Allen, P.~Swieskowski, and M.~Welsh.
\newblock {MoteLab: a Wireless Sensor Network Testbed}.
\newblock In {\em Proceedings of the 4th International Symposium on Information
  Processing in Sensor Networks (IPSN)}, 2005.

\bibitem{white:osdi:2002}
B.~White, J.~Lepreau, L.~Stoller, R.~Ricci, S.~Guruprasad, M.~Newbold,
  M.~Hibler, C.~Barb, and A.~Joglekar.
\newblock {An Integrated Experimental Environment for Distributed Systems and
  Networks}.
\newblock In {\em Proceedings of the Fifth Symposium on Operating Systems
  Design and Implementation (OSDI)}, 2002.

\bibitem{falcon-mobisys12}
T.~Yan, D.~Chu, D.~Ganesan, A.~Kansal, and J.~Liu.
\newblock Fast app launching for mobile devices using predictive user context.
\newblock In {\em Proceedings of the 10th international conference on Mobile
  systems, applications, and services}, MobiSys '12, pages 113--126, New York,
  NY, USA, 2012. ACM.

\end{thebibliography}
